\section{Introduction}

% \begin{frame}{Objectives}
% %\underline{\textsc{Some text:}}
% % \begin{small}
% % This is some small Text. 
% % \end{small}

% % \metroset{block=fill}
% % \begin{exampleblock}{\textsc{Example block}}
% % \begin{itemize}
% %     \item You know how to do itemize
% %     \item Also here
% % \end{itemize}
% % \end{exampleblock}
% \end{frame}

%1
\begin{frame}[fragile]{Introduction to statistics and data science}
\begin{block}{Why this course ?}
\begin{itemize}
    \item<+-> Data are omnipresent in engineering and business intelligence.
    \item<+-> Understanding them can often be challenging.
    \item<+-> Statistics provides the data analyst with a rigorous framework.
\end{itemize}
\end{block}
\begin{block}{Course contents:}
\begin{itemize}
    \item<+-> Concepts from probability theory.
    \item<+-> Visual data exploration.
    \item<+-> Inferential statistics.
    \item<+-> Hypothesis testing.
\end{itemize}
\end{block}
\end{frame}

%2
\begin{frame}[fragile]{Work organization}
\begin{itemize}
    \item<+-> All lectures will include course material and exercises.
    \item<+-> The notions are illustrated using real data and statistical software.
    \item<+->Should you encounter any difficulties, please do not hesitate to inquire.
    \item<+->Personal work is essential and must be completed regularly.
\end{itemize}
\end{frame}

%3
\begin{frame}[fragile]{Software}
\begin{itemize}
    \item<+-> Two software products will be utilized during the course:
    \begin{itemize}
        \item<+-> R, that is widely recognized as the prevailing standard within the statistical community.
        \item<+-> PowerBI, a tool that is being used with increasing frequency for the collection, processing, and visualization of data. It integrates almost seamlessly with R. 
    \end{itemize}
    \item<+-> Advanced concepts will not be covered. 
    \item<+-> Feel free to install both on your personal computer. Power BI Desktop cannot share dashboards
     unless you have a paid subscription, but this is not a significant issue when learning how to use the tool. 
    \end{itemize}
\end{frame}